---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metaconfoundr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/metaconfoundr)](https://CRAN.R-project.org/package=metaconfoundr)
[![R build status](https://github.com/malcolmbarrett/metaconfoundr/workflows/R-CMD-check/badge.svg)](https://github.com/malcolmbarrett/metaconfoundr/actions)
<!-- badges: end -->

The goal of metaconfoundr is to ...

## Installation

You can install the development version of metaconfoundr from GitHub with:

``` r
# if needed
# install.packages("remotes")
remotes::install_github("malcolmbarrett/metaconfoundr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(metaconfoundr)
## basic example code
```

Please note that the 'metaconfoundr' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
